Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-16 papers/thesis-kirstie/thesis.tex
edits on intro

Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
with one another like billiard balls. In the simplest case, the spheres are treated as having 
only a replulsive potential. This model, called the Hard-Sphere fluid, has served as the primary 
reference system for fluids. This is due to the fact that the repulsion between atoms is much 
stronger than their attraction, and so the repulsive potential, which is descriptive of 
hard spheres, dominates the interactions between atoms.

%Van der Waal took into account both the attraction, and the respulsion 
%of the impenetrable spheres in devopling his equation of state for an ideal gas in 1xxx.

In the late 1950s, Scaled Particle Theory (SPT) was developed~\cite{ReissSPT} from which an equation of state, and
an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be obtained~\cite{Hansen}. 
%Percus and Yevick took another approach to the Hard-Sphere fluid.
In 1961 Morita and Hiroike introduced a density functional approach to finding an expression for the 
excess Helmholtz free energy of inhomogeneous fluids~\cite{MoritaDFT}.
% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
In 1967 %the late 1960s, and early 1970s, 
Barker and Henderson~\cite{barker1967perturbation} adapted perturbation methods to use with fluids composed of 
spheres with softly fading potential functions rather than the hard edges of the spheres 
in the Hard-Sphere fluid. This was done by making the radius of the spheres temperature dependent, 
so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other when they collide at higher temperatures.
%The different size radii coorespond to the different depths to which two soft spheres would penetrate into each other when they collide. 
This reflects the idea that soft-spheres penetrate more deeply into each other when they collide at higher temperatures. 
Weeks, Chandler, and Anderson studied the soft repulsive portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which became known as the Weeks-Chandler-Anderson model fluid.

DFT Evans 1979

In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} which provides an expression for the excess 
Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
Tarazona later improved on the method by introducing a tensor version of the functional~\cite{tarazonaphi3}. 
In 1999 Schmidt adopted FMT for soft spheres forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. In this paper, 
Density Functional Theory will be used to predict the freezing of a Weeks-Chandler-Anderson model fluid 
by applying SFMT with Schmidt's proposed erf model\cite{schmidt2000fluid}. %for generating the weight funcions, 
%and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.
%\[{}\]
\newline
\newline
\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)

>>>>

Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
with one another like billiard balls. In the simplest case, the spheres are treated as having 
only a replulsive potential. This model, called the Hard-Sphere fluid, has served as the primary 
reference system for fluids. This is due to the fact that the repulsion between atoms is much 
stronger than their attraction, and so the repulsive potential, which is descriptive of 
hard spheres, dominates the interactions between atoms. 
In 1954 Zwanzig introduced the Square-well fluid which is a perturbation of the Hard-Sphere fluid 
created by adding a hard-edge attractive potential to the repulsive potential of the Hard-Sphere 
fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied spheres with softly fading replusive potentials which he 
modeled by making the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into each other 
when they collide at higher temperatures. 
%At higher temperatures the spheres would penetrate more deeply into each other when they collide, 
%and thus be represented by spheres with smaller radii. 
In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres with 
soft repulsive potentials to create the Barker-Henderson model fluid~\cite{barker1967perturbation}, 
where a soft-sphere fluid by is modeled by using a Hard-Sphere fluid with hard spheres of 
larger radii at low temperatures, and smaller radii at high temperatures. 
%
%an expression for the excess Helmholtz free energy of inhomogeneous fluids.
% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
%Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
%
%In 1967 Barker and Henderson~\cite{barker1967perturbation} adapted the perturbation method of Zwanzig 
%to use with fluids composed of spheres with softly fading potentials rather than the hard edges of the spheres 
%in the Hard-Sphere fluid. 
%This was done by making the radius of the spheres temperature dependent, 
%so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
%and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
%
%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other 
%when they collide at higher temperatures.
%The different size radii coorespond to the different depths to which two soft 
%spheres would penetrate into each other when they collide. 
%This reflects the idea that soft-spheres penetrate more deeply into each other when 
%they collide at higher temperatures. 
In 1971 Weeks, Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive portion of the 
Leonard-Jones potential~\cite{andersen1971relationship}, which became known as the 
Weeks-Chandler-Anderson model fluid.

While models for particular fluids were being created, powerful methods to describe the properties 
of fluids in general were also being developed. In the late 1950s, Scaled Particle 
Theory (SPT)~\cite{ReissSPT} was introduced from which an equation of state, and
an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
obtained~\cite{Hansen}. 
%Percus and Yevick took another approach to the Hard-Sphere fluid.
In 1961 Morita and Hiroike introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy, 
or the equilibrium Helmholtz free energy in the case of a canonical ensemble, and the cooresponding equilibrium 
number density profile for inhomogeneous fluids. To use this approach, a functional must be 
constructed for a particular fluid.
%DFT Evans 1979
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
to construct a functional for 
%which provides an expression for the excess Helmholtz free energy of an inhomogeneous 
the Hard-Sphere fluid. 
Tarazona later improved on the method by introducing a tensor version of the 
functional~\cite{tarazonaphi3}, and in 1999 Schmidt adopted FMT for soft spheres 
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

In this paper, Density Functional Theory (DFT) will be used to predict the freezing of a 
Weeks-Chandler-Anderson model fluid by applying SFMT with Schmidt's proposed 
erf model\cite{schmidt2000fluid} to form the functional. %for generating the weight funcions, 
%and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.
%\[{}\]
\newline
\newline
\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)
